Spin canonical divisors of log stacky curves

### 2015-06-20 stacky_write_up/spin_cring.tex
finished proof of complicated induction

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx',L'),(\sx,L)$ be tame, separably rooted log stacky spin curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose there is a birational map $\sx \rightarrow \sx'$, and let $J \subset \{1,
\ldots, r\}$. Suppose $e_i'+ 2\chi_J(i) = e_i$ where
$$
\chi_J(i) = \begin{cases}
	1, &\text{ if }i \in J\\
	0, &\text{ otherwise. } 
\end{cases}$$
Let $R'$ be the canonical ring associated to $\sx'$. Say $(\sx',J)$, {\bf admissible} if the following three conditions hold.
\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} \in R'$ so that $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
\end{align*}
so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
\item[(Ad-ii)] For all $i \in J$ and $z \neq y_{i,e_i'}$ a generator of $R'$ we have
\begin{align*}
	\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
\end{align*}
\item[(Ad-iii)] For all $i \in J$ we have
\begin{align*}
	\deg \lfloor e_i L \rfloor \geq \# S(i)
\end{align*}
where
\begin{align*}
	\eta(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{thm}
\label{thm:ramification_order_induction}
Suppose $(\sx',L')$ is a tame separably rooted log stacky spin curve is admissible with a generators $y_{i,e_i'} \in R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in H^0(\sx, e_i(K_\sx))$ so that
	\begin{align*}
	-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
\end{align*}
and
\begin{align*}
	\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-\frac{1}{\deg(y_{i,e_i'})e_j'}
\end{align*}
for all $j \in J$ with $j \neq i$.
\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0, b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally generate $R$ over $R'$.
\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with block order. Let $R = k[y,x]/I.$ Then,
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z \neq y_{i,e_i},y_{i,e_i'} \rangle 
\end{align*}
	as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by quadratics and that for all $i \in J$, we have $e_i > \deg z$ for any generator $z$ of $R'.$ Then any set of minimal generators for $I'$ together with any set of relations as in the previous part \todo{reference previous part as part c} minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}
\begin{proof}
Let $(\sx,L)$ and $(\sx',L')$ be the two log stacky spin curves in question, with coarse spaces $X = X'$. Define
\begin{align*}
	S(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
Define
\begin{align*}
	E_i = \sum_{j \in S(i)}^{}Q_j
\end{align*}
Note that
\begin{align*}
	\lfloor e_i L' \rfloor + Q_i \leq \lfloor e_i L \rfloor 
\end{align*}
and so we obtain an inclusion
\begin{align*}
	H^0(\sx',e_iL'-E_i + Q_j) \rightarrow H^0(\sx,e_iL - E_i) \subset H^0(\sx,e_iL).
\end{align*}
The assumption (Ad-iii) implies
\begin{align*}
	\deg \left( e_i L' - E_i \right) \geq 0
\end{align*}
implying that $H^0(\sx',e_iL'-E_i + Q_i)$ is base point free by Riemann Roch, and so a general element
\begin{align*}
	y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)
\end{align*}
satisfies
\begin{align*}
	-\ord_{Q_i}(y_{i,e_i}) = \lfloor e_i \subhalf {e_i'} \rfloor +1 = \frac{e_i - 1}{2}
\end{align*}
Therefore, $y_{i,e_i}$ satisfies the first part of the claim of $(a)$.

We next show $y_{i,e_i}$ also satisfies the second part of the claim of $(a),$ by considering seperately the cases of whether $j \in S(i),$ and $j \notin S(i)$.

If $j \in S(i)$ then, since $E_i \geq Q_j$, and $y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)$,
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq e_i\subhalf {e_j'} - 1 \leq e_i \subhalf{e_j'} - \frac{1}{e_j'}.
\end{align*}

If instead $j \notin S(i)$, then since $e_j' \nmid e_i$, we know $e_i\subhalf{e_j'} \notin \BZ$, so
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq \lfloor  e_i\subhalf{e_j'} \rfloor \leq e_i\subhalf{e_j'} - \frac{1}{e_j'},
\end{align*}
completing the proof of (a).

@@

To show (c), we wish to show $y_{i,e_i}z \in R'$ for $z \neq y_{i,e_i},y_{i,e_i'}$. Note that $f\in R$ lies in $R'$ if and only if for all $j \in J$ we have
\begin{align*}
	-\ord_{Q_j}(f) \leq \deg f \left( \subhalf {e_j'} \right) 
\end{align*}

>>>>

\begin{lem}
\label{lem:minimal_quadratic}

\end{lem}
\begin{proof}

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx',L'),(\sx,L)$ be tame, separably rooted log stacky spin curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose there is a birational map $\sx \rightarrow \sx'$, and let $J \subset \{1,
\ldots, r\}$. Suppose $e_i'+ 2\chi_J(i) = e_i$ where
$$
\chi_J(i) = \begin{cases}
	1, &\text{ if }i \in J\\
	0, &\text{ otherwise. } 
\end{cases}$$
Let $R'$ be the canonical ring associated to $\sx'$. Say $(\sx',J)$, {\bf admissible} if the following three conditions hold.
\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} \in R'$ so that $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
\end{align*}
so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
\item[(Ad-ii)] For all $i \in J$ and $z \neq y_{i,e_i'}$ a generator of $R'$ we have
\begin{align*}
	\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
\end{align*}
\item[(Ad-iii)] For all $i \in J$ we have for all $d \geq 0$,
\begin{align*}
	\deg \lfloor e_i L \rfloor \geq \# S(i,d)
\end{align*}
where
\begin{align*}
	S(i,d) = \{j \in J : j \neq i \text{ and }e_j'+2d \mid e_i+2d\}
\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{thm}
\label{thm:ramification_order_induction}
Suppose $(\sx',L')$ is a tame separably rooted log stacky spin curve is admissible with a generators $y_{i,e_i'} \in R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in H^0(\sx, e_i(K_\sx))$ so that
	\begin{align*}
	-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
\end{align*}
and
\begin{align*}
	\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-\frac{1}{\deg(y_{i,e_i'})e_j'}
\end{align*}
for all $j \in J$ with $j \neq i$.
\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0, b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally generate $R$ over $R'$.
\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with block order. Let $R = k[y,x]/I.$ Then,
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z \neq y_{i,e_i},y_{i,e_i'} \rangle 
\end{align*}
	as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by products of two monomials and that for all $i \in J$, we have $e_i > \deg z$ for any generator $z$ of $R'.$ Then any set of minimal generators for $I'$ together with any set of relations as in part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}
\begin{proof}
Let $(\sx,L)$ and $(\sx',L')$ be the two log stacky spin curves in question, with coarse spaces $X = X'$. Define
\begin{align*}
	S(i,0) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
Define
\begin{align*}
	E_i = \sum_{j \in S(i,0)}^{}Q_j
\end{align*}
Note that
\begin{align*}
	\lfloor e_i L' \rfloor + Q_i \leq \lfloor e_i L \rfloor 
\end{align*}
and so we obtain an inclusion
\begin{align*}
	H^0(\sx',e_iL'-E_i + Q_j) \rightarrow H^0(\sx,e_iL - E_i) \subset H^0(\sx,e_iL).
\end{align*}
The assumption (Ad-iii) implies
\begin{align*}
	\deg \left( e_i L' - E_i \right) \geq 0
\end{align*}
implying that $H^0(\sx',e_iL'-E_i + Q_i)$ is base point free by Riemann Roch, and so a general element
\begin{align*}
	y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)
\end{align*}
satisfies
\begin{align*}
	-\ord_{Q_i}(y_{i,e_i}) = \lfloor e_i \subhalf {e_i'} \rfloor +1 = \frac{e_i - 1}{2}
\end{align*}
Therefore, $y_{i,e_i}$ satisfies the first part of the claim of $(a)$.

We next show $y_{i,e_i}$ also satisfies the second part of the claim of $(a),$ by considering seperately the cases of whether $j \in S(i,0),$ and $j \notin S(i,0)$.

If $j \in S(i,0)$ then, since $E_i \geq Q_j$, and $y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)$,
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq e_i\subhalf {e_j'} - 1 \leq e_i \subhalf{e_j'} - \frac{1}{e_j'}.
\end{align*}

If instead $j \notin S(i,0)$, then since $e_j' \nmid e_i$, we know $e_i\subhalf{e_j'} \notin \BZ$, so
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq \lfloor  e_i\subhalf{e_j'} \rfloor \leq e_i\subhalf{e_j'} - \frac{1}{e_j'},
\end{align*}
completing the proof of (a).

@@

To show (c), we wish to show $y_{i,e_i}z \in R'$ for $z \neq y_{i,e_i},y_{i,e_i'}$. Note that $f\in R$ lies in $R'$ if and only if for all $j \in J$ we have
\begin{align*}
	-\ord_{Q_j}(f) \leq \deg f \left( \subhalf {e_j'} \right) 
\end{align*}
We now check this in three cases, depending on whether $j \notin \{i \} \cup S(i,0),j = i,$ or $j \in S(i,0)$.

\vskip.1in
\noindent
{\sf Case 1: $j \notin \{i \} \cup S(i,0).$}

Here, $-\ord_{Q_j}(L) = -\ord_{Q_j}(L'),$ so
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z) \leq e_i \subhalf {e_j'}+ \deg z \subhalf {e_j'} = \deg (y_{i,e_i}z) \subhalf{e_j'}. 
\end{align*}

\vskip.1in
\noindent
{\sf Case 2: $j =i.$}

By part (a), condition (Ad-ii), and Lemma ~\ref{lem:admissible_inequality} we have
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z) &\leq \frac{e_i-1}{2} + \deg z \left( \subhalf{e_i'} \right) -\frac{1}{e_i'}
	& = \frac{e_i-1}{2}-\frac{1}{e_i'}+\left( \subhalf{e_i'} \right)\\
	&= \frac{e_i(e_i-3)}{2(e_i-2)}+\left( \subhalf{e_i'} \right)\\
	&= \deg (y_{i,e_i}z) \subhalf{e_j'}.
\end{align*}

\vskip.1in
\noindent
{\sf Case 3: $j \in S(i,0).$}
In this case, we similarly obtain
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z) &\leq  e_i \subhalf{e_j'} +\deg z \subhalf{e_j'} = \deg (y_{i,e_i}z) \subhalf{e_j'},
\end{align*}
completing part (c).

Next, note that the minimal generators of $I'$ together with the new generators of $I$ given in (c) have independent initial terms by construction, and therefore are minimal by Lemma ~\ref{lem:minimal_quadratic}.

Finally, to check (e), we will first show that $(\sx,J)$ is admissible. We know (Ad-i) is satisfied by part (b), taking the $y_{i,e_i}$ as the generators in degree $e_i$. Next, (Ad-ii) is strictly monotonic in the $e_i$, and hence also holds for $(\sx,J)$. Finally, if (Ad-iii) holds for $e$ then it holds for $e+2$ by definition. This is where we use that (Ad-iii) holds for $d > 0$ and not just for $d = 0$.

Finally, if $(\sx,J)$ then so is $(\sx,W)$ for any $W \subset J$, since each of the conditions (Ad-i), (Ad-ii), and (Ad-iii) hold for $W$ if they hold for $J$.
\end{proof}
